Peer-graded Assignment: Course Project 1
=========================================


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#load the data
```{r loaddata}
activity<- read.table(unz("activity.zip","activity.csv"), header = TRUE, sep = ",")
head(activity)
```
#process the data
```{r process data}
activity$date <- as.Date(activity$date, format = "%Y-%m-%d")
```

# 1) What is mean total number of steps taken per day?
## calculate the total step number per day

```{r}
sum_day <- data.frame(tapply(activity$steps, as.factor(activity$date), sum, na.rm = TRUE))
colnames(sum_day) <- c("steps")
```

## Calculate and report the mean and median of the total number of steps taken per day.

```{r}
mean(sum_day$steps)
```
```{r}
median(sum_day$steps)
```

# 2) What is the average daily activity pattern?

##Make a time series plot (i.e. type = "l") of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all days (y-axis).

```{r}
avg_steps <- aggregate(steps~interval, data = activity, mean, na.rm = TRUE)
plot(avg_steps$interval, avg_steps$steps, type = "l", lwd = 2, col = "navy",
     main = "Time series: average number of steps taken",
     xlab = "5-minute interval", ylab = "Average number of steps")



```


##Which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps?

```{r}
avg_steps$interval[which.max(avg_steps$steps)]
```

The 835-th 5-minute interval contains the maximum number of steps.

# 3) Imputing missing values

## calculate and report the number of missing values in the dataset

```{r}
sum(is.na(activity))
```

## devise a strategy for filling in all of missing values in the dataset
```{r}
activity_new <- activity
for (i in avg_steps$interval) {
        activity_new[activity_new$interval == i & is.na(activity_new$steps),]$steps <-
                avg_steps$steps[avg_steps$interval == i]
}
head(activity_new)
```
## Make a histogram of the total number of steps taken each day and Calculate and report the mean and median total number of steps taken per day. Do these values differ from the estimates from the first part of the assignment? What is the impact of imputing missing data on the estimates of the total daily number of steps?
```{r}
total_step_new <- aggregate(steps~date, data = activity_new, sum, na.rm = TRUE)
hist(total_step_new$steps, breaks = 20, 
     main = "Total number of steps taken each day (imputed)",
     col = "grey", border = "white", xlab = "Step")
```

```{r}
mean(total_step_new$steps)
```
```{r}
median(total_step_new$steps)
```

# 4) Are there differences in activity patterns between weekdays and weekends?

##Create a new factor variable in the dataset with two levels – “weekday” and “weekend” indicating whether a given date is a weekday or weekend day.

```{r}
activity_new$day <- weekdays(activity_new$date)
activity_new$week <- ""
activity_new[activity_new$day == "Saturday" | activity_new$day == "Sunday", ]$week <-"weekend"
activity_new[!(activity_new$day == "Saturday" | activity_new$day == "Sunday"),]$week <- "weekday"
activity_new$week <- factor(activity_new$week)
```

## Make a panel plot containing a time series plot (i.e. type = "l") of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all weekday days or weekend days (y-axis). See the README file in the GitHub repository to see an example of what this plot should look like using simulated data.
```{r}
library(ggplot2)
avg_steps_day_type <- aggregate(steps ~ interval+ week, data = activity_new, mean)
plt <- ggplot(avg_steps_day_type, aes(interval, steps))+
        geom_line(stat = "identity", aes(color = week))+
        theme_gray()+
        facet_grid(week~.,scales = "fixed", space = "fixed")+
        labs(x = "Interval", y = "Steps")+
        ggtitle("Steps per interval by day type")


print(plt)
```

The users start a bit later on weekend and have less steps in weekend morning compared to weekdays.










